Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-02-04 2_SSL_EES/article/article_main.tex
finalized figure captions

\begin{figure*}[h]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/history_efficacy.pdf}
 \caption{Historical development of the luminous efficacy ($K$) of the most widely-used lighting technologies in human history from 1850 to 2020. Data shown for best performers by year of market introduction, with dashed lines. Dashed lines indicate the average improvement calculated from a 3rd-order polynomial fit to the data. The physical limit for an ideal light source with a colour rendering index of CRI=90, denoted as $K_max^{CRI90}$, is shown as a black horizontal line, as per calculations by Murphy et al. \cite{Murphy2012}. The magnified plot shows the progress in cool white LEDs from 1996 to 2020, with the dashed line indicating a linear rate of efficacy improvement of 10lm/W per year. For comparison, efficacies of best performers in legacy lighting technologies for 2020 are shown as coloured horizontal lines. Note the logarithmic scale of the vertical axis on the main plot and the linear scale on the magnified plot. Abbreviations: HID - High-Intensity Discharge; CFL - Compact Fluorescent Lamp; Hal. - Halogen, Incd. - Incandescent. Source: own synthesis of published data based on a visual approach proposed by Azevedo et al. \cite{azevedo2009transition}. See Supplementary Information, Section 5 (Complete List of Sources for Figures), for the full list of sources and references.}
 \label{fgr:history_efficacy}
\end{figure*}

@@

\begin{figure*}[h!]
 \centering
 \includegraphics[width=17.5cm]{2_SSL_EES/article/figures/subefficiencies_progress.pdf}
 \caption{Historical improvements in light-emitting diode technology. Panels A1-A2 show they key metrics used to compute forward voltage efficiency $\eta_{Vf}$ and droop efficiency $\eta_{droop}$. Panels B1-B2 show the sub-efficiencies internal quantum efficiency $\eta_{IQ}$ and light extraction efficiency $\eta_{LE}$. \newline
 \textbf{Panel A1:} Device forward voltage at a test current of I=350mA. The physical limit for a blue light wavelength of 450nm without electric pumping is shown for reference. Data points for devices released in 2020 by various manufacturers are shown in an inset plot.
 \textbf{Panel A2:} Efficiency droop at the test current of I=350mA.
 \textbf{Panel B1:} Internal quantum efficiency, for different chip architectures, by type of measurement used. Note that the artifact in the DOE Average around 2013 is due to a change in definition for internal quantum efficiency laid out in \cite{doe_ssl_multiyear_2013}. Measurement methods: PL - Photoluminescence\cite{Shim_2018}, EL - Electroluminescence\cite{Getty_2009};  ABC Model\cite{Karpov_2014}, Ideality Factor\cite{Masui_2010}. “Not reported” denotes data points where the IQE measurement technique was not reported. Whiskers indicate reported ranges, where applicable.
 \textbf{Panel B2:} Light extraction efficiency for different chip architectures, by type of specific technology used to improve light-extraction efficiency. Abbreviations: TFFC - Thin-Film Flip-Chip; FC - Flip-Chip; CSP - Chip-Scale Package; ITO - Indium Tin Oxide; ZnO - Zinc Oxide; PSS - Patterned Sapphire Substrate; electr. - electrode; text. - textured.
 Source: own synthesis of data from the full list of sources provided in the Supplementary Information, Section 5. Data on average device performance adapted from U.S. Department of Energy (DOE) Reports\cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_rnd_2016}\cite{doe_ssl_rnd_2018} and an Osram company report\cite{beale_leds_2015}.}
 \label{fgr:subeff}
\end{figure*}

\begin{figure*}[h!]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/breakthroughs_efficiency.pdf}
 \caption{Contribution of innovations and technology spillovers to the historical progress in sub-efficiencies of phosphor-converted warm white LEDs with test currents of at least 350mA. Vertical bars represent LED technology innovations, with purple bars indicating innovations driven by technology spillovers (annotated and listed in an inset table) and grey bars indicating all other improvements identified in this study. Horizontal coloured lines indicate state-of-the-art sub-efficiency levels for five years: 1997, 2003, 2010, 2016, and 2020. Physical limits on sub-efficiencies are indicated by black horizontal lines. The overall LED lamp efficiency, $\eta_L$, is displayed in the rightmost column. $\eta_Vf$ - forward voltage efficiency; $\eta_IQ$ - internal quantum efficiency; $\eta_Droop$ - efficiency droop; $\eta_LE$ - light extraction efficiency; $\eta_(C_ER)$ - conversion efficiency for red phosphors; $\eta_(C_(EY/G))$ - conversion efficiency for yellow/green phosphors; $\eta_S$ - spectral efficiency; PSS - patterned sapphire substrate, ITO - indium tin oxide current spreading layer; 258, SLA, PFS - different phosphors used in white LEDs, see Table 3. "N/A" denotes sub-efficiencies that could not be calculated for 1997 due to $\eta_Vf$, $\eta_droop$ depending on current, which was below 350mA at the time, and warm white spectrum LEDs not available at the time for $\eta_(C_ER)$,$\eta_(C_(EY/G))$  and $\eta_S$ The physical limit on $\eta_Vf$ exceeds 100\% and depends on electrical device parameters \cite{david2016electrical}. Source: own elaboration based on data in Figure \ref{fgr:waterfall} and Figure \ref{fgr:subeff}, as detailed in Section \ref{sec:methods}.}
 \label{fgr:breakthroughs_efficiency}
\end{figure*}

\begin{figure*}[h!]
 \centering
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2003.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2010.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2016.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2020.pdf}
 \caption{Waterfall diagrams of the loss channels in a generic mid/high-power LED package for 2003, 2010, 2016 and 2020 (top to bottom panels), normalized to 1 Watt of electric power input (yellow bar on the left). Sub-efficiencies corresponding to each loss channel are listed below each column and described in Section \ref{sec:device_performance_metrics}. Numbers for each loss channel indicate energy losses both in relative terms of input power (in percent) at the point of the channel and absolute values (in Watts). Percentages for red, green and blue loss channels indicate losses of remaining red/green/blue light energy. The corresponding LED architectures from Figure \ref{fgr:chip_architecture_overview} are shown for reference: 2003 (top panel) - flip-chip with YGAG phosphor; 2010 (bottom panel) - flip-chip with 258 phosphor; 2016 (top panel) - flip-chip with PFS phosphor; 2020 (bottom panel) - flip-chip with SALON phosphor. Details on phosphors are provided in Table \ref{tab:phosphors} and SI section XX. Overall LED package efficiency is $\eta_L = 5.8\%$ in 2003 and $\eta_L = 12.7\%$ in 2010. Abbreviations: Scat. = Scattering. Source (Efficiency): own elaboration based on data in Figure \ref{fgr:subeff}, as detailed in Section \ref{sec:methods}. Source (spectral data): Adapted from publications on the respective phosphors: YGAG (2003)\cite{Mueller2002}, 258 (2010)\cite{MuellerMach2005}, PFS (2016)\cite{Murphy2015}, SALON (2019)\cite{Hoerder2019}.}
 \label{fgr:waterfall}
\end{figure*}

@@

\begin{table*}[h!]
    \small
    \centering
    \begin{tabularx}{\textwidth}{|l|l|l|X|X|l|}
    \hline
        \textbf{Year} & \textbf{Desig.} & \textbf{Chemical formula} & \textbf{Description} & \textbf{Significance} & \textbf{SI Sec.} \\ \hline
        1996 & YAG & $Y_3 Al_5 O_{12}:Ce$ & Yttrium aluminium garnet (YAG) phosphor activated with cerium & First LED phosphor, enabled white LEDs & SI 3 \\ \hline
        1996 & YGAG & $(Y_{1-x} Gd_x)_3 Al_5 O_{12}:Ce$ & Gadolinium-doped YAG phosphor & First red-shifted phosphor, enabled warm white LEDs & SI 3.1 \\ \hline
        2002 & 258 & $(Ba,Sr)_2 Si_5 N_8:Eu^{2+}$ & Europium-doped nitridosilicate phosphor & First red LED phosphor & SI 3 \\ \hline
        2003 & QD & N/A & Quantum dot-based phosphor & First use of QD for LED light down conversion & SI 3 \\ \hline
        2005 & PFS & $K_2 SiF_6: Mn^{4+}$ & Manganese-activated potassium fluorosilicate (PFS) phosphor & First ultra-narrow-band red LED phosphor & SI 3 \\ \hline
        2013 & SLA & $Sr[Li Al]_3 N_4 ]:Eu^{2+}$ & Europium-doped cuboidal nitridolithoaluminate phosphor & Improved narrow-band red phosphor & SI 3 \\ \hline
        2016 & SALON & $Sr[Li_2 Al_2 O_2 N_2]:Eu^{2+}$ & Europium-doped oxonitride phosphor & High-performance ultra-narrow-band red phosphor & SI 3 \\ \hline
    \end{tabularx}
    \caption{LED light down conversion inventions related to improvements in consumer experience metrics identified in this study. Note: Year column represents the earliest reported application of invention in white LEDs. These years are different from the years in Figure \ref{fgr:consumer_experience} that correspond to the earliest publication of spectral data for representative LED products  that relied on those inventions. Section in SI refers to a section in the Supplementary Information that contains references and a detailed description of the history of corresponding invention. Abbreviations: Desig. - Designation, Sec. - Section.}
    \label{tab:phosphors}
\end{table*}

\begin{figure*}[h!]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/breakthroughs_consumer-experience.pdf}
 \caption{Historical improvements in consumer experience metrics of phosphor-converted white LEDs. Circle dots show data for the earliest identified representative white LED products with published spectral data that used phosphors listed in Table 3, each indicated by the phosphor label and publication year. Corresponding spectral data used to calculate colour temperature values is provided in Figure SI1 in SI section 2. Three data points representing compact fluorescent lamps (CFLs)\cite{cie_reference}, shown as black triangles, are provided for comparison. Horizontal lines represent typical colour temperatures of “traditional” light sources, shown for reference. Desirable range of colour temperatures for home illumination, indicated by a vertical red arrow, lies between two horizontal orange lines representing typical incandescent light and warm daylight colour temperatures. Horizontal red arrow indicates desirable higher values of colour rendering index (CRI). LED products based on the following phosphor innovations from the following LED manufacturers are represented : YAG - Nichia, 1998 \cite{bando1998development}; YGAG - Lumileds, 2002 \cite{Mueller2002}; 258 - Lumileds, 2005 \cite{MuellerMach2005}; SLA - Lumileds, 2014 \cite{Pust2014}; PFS - GE, 2015 \cite{Murphy2015}; QD - Lumileds, 2017 \cite{lumileds2016qd}, Osram, 2019 \cite{osram2019qd}; SALON - Osram, 2019 \cite{Hoerder2019}.}
 \label{fgr:consumer_experience}
\end{figure*}

@@

\begin{figure*}[h!]
 \centering
 \includegraphics[width=17.5cm]{2_SSL_EES/article/figures/costmodel_results_years.pdf}
 \caption{Manufacturing cost structure modelled for a single low-to-mid power GaN, classic chip, phosphor-converted white LED package, assuming an ideal factory with state-of-the-art equipment at a U.S. location. Panels A-C: Waterfall diagrams of LED manufacturing cost split by manufacturing process steps for years 2003, 2012 and 2020. Process steps on the horizontal axis are sequenced from left to right in the same order as in the modelled LED manufacturing process. Panel F: Cumulative manufacturing yield after each process step for years 2003, 2012 and 2020. Abbreviations: Litho – Lithographic Process, Insp. – Inspection, Depn. – Deposition, CMP - Chemical-Mechanical Planarization}
 \label{fgr:costmodel}
\end{figure*}

>>>>

\begin{figure*}[h]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/history_efficacy.pdf}
 \caption{Historical development of the luminous efficacy ($K$) of the most widely-used lighting technologies in human history. Data points indicate best performers by year of market introduction. Dashed lines indicate an average improvement for each technology, computed from a 3rd-order polynomial fit to the data. The physical limit for an ideal light source with a colour rendering index of CRI=90, denoted as $K_max^{CRI90}$, is shown as a black horizontal line, as per calculations by Murphy et al. \cite{Murphy2012}. The magnified plot shows the progress in cool white LEDs from 1996 to 2020, with the dashed line indicating a linear rate of efficacy improvement of 10lm/W per year. For comparison, efficacies of best performers in legacy lighting technologies for 2020 are shown as coloured horizontal lines. Note the logarithmic scale of the vertical axis on the main plot and the linear scale on the magnified plot. Abbreviations: HID - High-Intensity Discharge; CFL - Compact Fluorescent Lamp; Hal. - Halogen, Incd. - Incandescent. Source: own synthesis of published data based on a visual approach proposed by Azevedo et al. \cite{azevedo2009transition}. See Supplementary Information, Section 5 (Complete List of Sources for Figures), for the full list of sources and references.}
 \label{fgr:history_efficacy}
\end{figure*}

@@

\begin{figure*}[h!]
 \centering
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/subefficiencies_progress.pdf}
 \caption{Historical improvements in light-emitting diode technology. Panels A1-A2 show they key metrics used to compute forward voltage efficiency $\eta_{Vf}$ and droop efficiency $\eta_{droop}$. Panels B1-B2 show the sub-efficiencies internal quantum efficiency $\eta_{IQ}$ and light extraction efficiency $\eta_{LE}$. \newline
 \textbf{Panel A1:} Device forward voltage at a test current of I=350mA. The physical limit for a blue light wavelength of 450nm without electric pumping is shown for reference. Data points for devices released in 2020 by various manufacturers are shown in an inset plot.
 \textbf{Panel A2:} Efficiency droop at the test current of I=350mA.
 \textbf{Panel B1:} Internal quantum efficiency, for different chip architectures, by type of measurement used. Note that the artifact in the DOE Average around 2013 is due to a change in definition for internal quantum efficiency laid out in \cite{doe_ssl_multiyear_2013}. Measurement methods: PL - Photoluminescence\cite{Shim_2018}, EL - Electroluminescence\cite{Getty_2009};  ABC Model\cite{Karpov_2014}, Ideality Factor\cite{Masui_2010}. “Not reported” denotes data points where the IQE measurement technique was not reported. Whiskers indicate reported ranges, where applicable.
 \textbf{Panel B2:} Light extraction efficiency for different chip architectures, by type of specific technology used to improve light-extraction efficiency. The source of data points is shown for reference: simulation - ray-tracing computer simulation; experimental (absolute) - light-extraction efficiency given directly; experimenal (relative) - relative improvement over baseline chip architecture given. Abbreviations: TFFC - Thin-Film Flip-Chip; FC - Flip-Chip; CSP - Chip-Scale Package; ITO - Indium Tin Oxide; ZnO - Zinc Oxide; PSS - Patterned Sapphire Substrate; electr. - electrode; text. - textured.
 Source: own synthesis of data from the full list of sources provided in the Supplementary Information, Section 5. Data on average device performance adapted from U.S. Department of Energy (DOE) Reports\cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_rnd_2016}\cite{doe_ssl_rnd_2018} and an Osram company report\cite{beale_leds_2015}.}
 \label{fgr:subeff}
\end{figure*}

\begin{figure*}[h!]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/breakthroughs_efficiency.pdf}
 \caption{Contribution of innovations and technology spillovers to the historical progress in sub-efficiencies of phosphor-converted warm white LEDs with test currents of at least 350mA. Vertical bars represent LED technology innovations, with purple bars indicating innovations driven by technology spillovers (annotated and listed in an inset table) and grey bars indicating all other improvements identified in this study. Horizontal coloured lines indicate state-of-the-art sub-efficiency levels for the four years used in Figure \ref{fgr:waterfall}: 2003, 2010, 2016, and 2020. To provide additional historical context, data for 1997 is included for selected sub-efficiencies. "N/A" denotes sub-efficiencies where 1997 data could not be calculated to different reasons: $\eta_{Vf}$, $\eta_{droop}$ depend on the device current, which was below 350mA in 1997, making comparison with contemporary devices difficult. $\eta_{C(R)}$,$\eta_{C(Y/G)}$ and $\eta_S$ are relevant only to warm white spectrum LEDs, which were not available in 1997. Physical limits on sub-efficiencies are indicated by black horizontal lines. The overall LED lamp efficiency, $\eta_L$, is displayed in the rightmost column. $\eta_{V_f}$ - forward voltage efficiency; $\eta_IQ$ - internal quantum efficiency; $\eta_Droop$ - efficiency droop; $\eta_{LE}$ - light extraction efficiency; $\eta_{C(R)}$ - conversion efficiency for red phosphors; $\eta_{C(Y/G)}$ - conversion efficiency for yellow/green phosphors; $\eta_S$ - spectral efficiency; PSS - patterned sapphire substrate, ITO - indium tin oxide current spreading layer; 258, SLA, PFS - different phosphors used in white LEDs, see Table \ref{tab:phosphors}.  Note that physical limit on $\eta_{V_f}$ is above 100\% due to quantum effects which depends on electrical device parameters \cite{david2016electrical}. Source: own elaboration based on data in Figure \ref{fgr:waterfall} and Figure \ref{fgr:subeff}, as detailed in Section \ref{sec:methods}.}
 \label{fgr:breakthroughs_efficiency}
\end{figure*}

\begin{figure*}[h!]
 \centering
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2003.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2010.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2016.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2020.pdf}
 \caption{Waterfall diagrams of the loss channels in a generic mid/high-power LED package for 2003, 2010, 2016 and 2020 (top to bottom panels), normalized to 1 Watt of electric power input (yellow bar on the left). Sub-efficiencies corresponding to each loss channel are listed below each column and described in Section \ref{sec:device_performance_metrics}. Numbers for each loss channel indicate energy losses both in relative terms of input power (in percent) at the point of the channel and absolute values (in Watts). Percentages for red, green and blue loss channels indicate losses of remaining red/green/blue light energy. The corresponding LED architectures and associated down-conversion phosphors are shown for reference: 2003 - flip-chip with YGAG phosphor; 2010 - flip-chip with 258 phosphor; 2016 - flip-chip with PFS phosphor; 2020 - flip-chip with SALON phosphor. A more complete overview of LED chip architectures is provided in Figure \ref{fgr:chip_architecture_overview}. Details on phosphor chemical composition are provided in Table \ref{tab:phosphors} with additional data on the spectral performance and development history collected in Section 4 in the supplementary information. Overall LED package efficiency is $\eta_L = 5.8\%$ in 2003, $\eta_L = 12.7\%$ in 2010, $\eta_L = 32.6\%$ in 2016 and $\eta_L = 38.8\%$ in 2010. Abbreviations: Scat. = Scattering. Source (Efficiency): own elaboration based on data in Figure \ref{fgr:subeff}, as detailed in Section \ref{sec:methods}. Source (spectral data): Adapted from publications on the respective phosphors: YGAG (2003)\cite{Mueller2002}, 258 (2010)\cite{MuellerMach2005}, PFS (2016)\cite{Murphy2015}, SALON (2019)\cite{Hoerder2019}.}
 \label{fgr:waterfall}
\end{figure*}

@@

\begin{table*}[h!]
    \small
    \centering
    \begin{tabularx}{\textwidth}{|l|l|l|X|X|l|}
    \hline
        \textbf{Year} & \textbf{Desig.} & \textbf{Chemical formula} & \textbf{Description} & \textbf{Significance} & \textbf{SI Sec.} \\ \hline
        1996 & YAG & $Y_3 Al_5 O_{12}:Ce$ & Yttrium aluminium garnet (YAG) phosphor activated with cerium & First LED phosphor, enabled white LEDs & SI 3 \\ \hline
        1996 & YGAG & $(Y_{1-x} Gd_x)_3 Al_5 O_{12}:Ce$ & Gadolinium-doped YAG phosphor & First red-shifted phosphor, enabled warm white LEDs & SI 3.1 \\ \hline
        2002 & 258 & $(Ba,Sr)_2 Si_5 N_8:Eu^{2+}$ & Europium-doped nitridosilicate phosphor & First red LED phosphor & SI 3 \\ \hline
        2003 & QD & N/A & Quantum dot-based phosphor & First use of QD for LED light down conversion & SI 3 \\ \hline
        2005 & PFS & $K_2 SiF_6: Mn^{4+}$ & Manganese-activated potassium fluorosilicate (PFS) phosphor & First ultra-narrow-band red LED phosphor & SI 3 \\ \hline
        2013 & SLA & $Sr[Li Al]_3 N_4 ]:Eu^{2+}$ & Europium-doped cuboidal nitridolithoaluminate phosphor & Improved narrow-band red phosphor & SI 3 \\ \hline
        2016 & SALON & $Sr[Li_2 Al_2 O_2 N_2]:Eu^{2+}$ & Europium-doped oxonitride phosphor & High-performance ultra-narrow-band red phosphor & SI 3 \\ \hline
    \end{tabularx}
    \caption{LED down-conversion materials (phosphors and quantum dots) related to improvements in consumer experience metrics identified in this study. The \textit{year} column represents the earliest reported application of invention in white LEDs. These differ from the years used in Figure \ref{fgr:consumer_experience} that correspond to the earliest publication of spectral data for representative LED products that relied on those materials. The \textit{SI Sec.} column refers to the corresponding section in the Supplementary Information that contains references and a detailed description of the history of corresponding invention. Abbreviations: Desig. - Designation, Sec. - Section.}
    \label{tab:phosphors}
\end{table*}

\begin{figure*}[h!]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/breakthroughs_consumer-experience.pdf}
 \caption{Historical improvements in consumer experience metrics of phosphor-converted white LEDs. Data points show the color temperature and colour rendering index performance of the earliest identified representative white LED products with published spectral data that used phosphors listed in Table \ref{tab:phosphors}, each indicated by the phosphor label and publication year. Corresponding spectral data used to calculate colour temperature values is provided in Figure SI 9 in section 4 of the supplementary information. Three data points representing compact fluorescent lamps (CFLs)\cite{cie_reference}, shown as black triangles, are provided for comparison. Horizontal lines represent typical colour temperatures of “traditional” light sources, shown for reference. The desirable range of colour temperatures for home illumination, indicated by a vertical red arrow, lies between two horizontal orange lines representing typical incandescent light and warm daylight colour temperatures. Horizontal red arrow indicates desirable higher values of colour rendering index (CRI). LED products based on the following phosphor innovations from the following LED manufacturers are represented : YAG - Nichia, 1998 \cite{bando1998development}; YGAG - Lumileds, 2002 \cite{Mueller2002}; 258 - Lumileds, 2005 \cite{MuellerMach2005}; SLA - Lumileds, 2014 \cite{Pust2014}; PFS - GE, 2015 \cite{Murphy2015}; QD - Lumileds, 2017 \cite{lumileds2016qd}, Osram, 2019 \cite{osram2019qd}; SALON - Osram, 2019 \cite{Hoerder2019}.}
 \label{fgr:consumer_experience}
\end{figure*}

@@

\begin{figure*}[h!]
 \centering
 \includegraphics[width=17.5cm]{2_SSL_EES/article/figures/costmodel_results_years.pdf}
 \caption{Manufacturing cost structure modelled for a single low-to-mid power GaN, classic chip, phosphor-converted white LED package, assuming an ideal factory with state-of-the-art equipment at a U.S. location.  Panels A-C: Waterfall diagrams of LED manufacturing cost split by manufacturing process steps for years 2003, 2012 and 2020. Process steps on the horizontal axis are sequenced from left to right in the same order as in the modelled LED manufacturing process. Panel F: Cumulative manufacturing yield after each process step for years 2003, 2012 and 2020. For a visual representation of the overview of the manufacturing process, compare the diagrams in Section 2 of the supplementary information. Abbreviations: Litho – Lithographic Process, Insp. – Inspection, Depn. – Deposition, CMP - Chemical-Mechanical Planarization}
 \label{fgr:costmodel}
\end{figure*}

### 2023-02-04 2_SSL_EES/article/article_supplementary.tex
finalized figure captions

@@

>>>>

\renewcommand{\figurename}{Figure SI}
\renewcommand{\tablename}{Table SI}

@@

\section{Identified LED Innovations}

\begin{table}[H]
    \begin{tabularx}{\textwidth}{ |l|X|X|l|l| }
        \multicolumn{5}{c}{Forward Voltage Efficiency (VFE) $\eta_{Vf}$} \\
        \hline
            \textit{Year} & \textit{Breakthrough} & \textit{Affects} & \textit{S/O} & \textit{Source} \\
        \hline
            1983 & Indium Tin Oxide \newline Current Spreading Layer & Contact Resistance & Yes (T,S) & \cite{margalith1999indium}\\
        \hline
            1998 & Digited Electrodes & Contact Resistance & Weak & \cite{steigerwald2001electrode} \\
        \hline
            ongoing & Epitaxy Improvements \newline and better Doping & Polarization and \newline Bulk Resistance & No & I \\
        \hline
        $\sim$2000 & Silver p-Contacts & Contact Resistance & No & \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Light Extraction Efficiency (LEE) $\eta_{LE}$} \\
        \hline
            \textit{Year} & \textit{Breakthrough} & \textit{Affects} & \textit{S/O} & \textit{Source} \\
        \hline
            $<$ 2003 & Optimization for  \newline Cavity Effects & Other & No & I, \cite{Shen2003} \\
        \hline
            (1973) 1993 & "Chip \newline Randomization" & Total Reflection \newline Absorption & No & I, \cite{bergh1973surface}\cite{Schnitzer1993} \\
        \hline
            1993 & Thin-Film \newline Chip Architecture & Absorption & No & I, \cite{Schnitzer1993} \\
        \hline
            $<2001$ & Patterned Sapphire \newline Substrate  & Total Reflection \newline Absorption & Yes (T,S) & I, \cite{Tadatomo2001} \\
        \hline
            ongoing & Chip Design for \newline high LEE & Total Reflection \newline Absorption & No & I, \cite{Haerle2004} \\
        \hline
            $\sim$2000 & Silver p-Contacts & Absorption & No & I, \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Internal Quantum Efficiency (IQE) $\eta_{IQ}$} \\
        \hline
            \textit{Year} & \textit{Breakthrough} & \textit{Affects} & \textit{S/O} & \textit{Source} \\
        \hline
            1994 & DH & $>$ Radiative Recombination \newline Probability & No & \cite{Nakamura1994} \\
        \hline
            1996 & MQW & $>$ Radiative Recombination \newline Probability & No & \cite{Koike1996} \\
        \hline
            ongoing & Active Region \newline Doping & Both & No & I, \cite{schubert2018light} \\
        \hline
            ongoing & Epitaxy \newline Improvements & Both & No & I \\
        \hline
            ongoing & Chip Architecture \newline Improvements & Both & No & I \\
        \hline
        \multicolumn{5}{c}{(Phosphor) Conversion Efficiency (CE) $\eta_{C}$} \\
        \hline
            \textit{Year} & \textit{Breakthrough} & \textit{Affects} & \textit{S/O} & \textit{Source} \\
        \hline
            ongoing & Lower Current Density & Current Density & No & I \\
        \hline
            ongoing & Epitaxy \newline Improvements & $\ominus/\oplus$ Distribution \newline Defect Density & No & I, \cite{bhardwaj2016progress} \\
        \hline
            ongoing & Chip Architecture \newline Improvements  & $\ominus/\oplus$ Distribution \newline Defect Density & No & I, \cite{Wildeson2017} \\
        \hline
            $<$ 2017 & Defect Getting \newline Underlayer & Defect Density & No & I, \cite{haller2017burying} \\
        \hline
        \end{tabularx}
    \caption{Technology breakthroughs and spillovers related to a subset of the sub-efficiencies considered. For a list of innovations in spectral efficiency, compare Table 1 in the main article. Listed are improvements in forward voltage efficiency, light extraction efficiency, internal quantum efficiency and conversion efficiency. No  Year column indicates first identified use in solid-state lighting. S/O column indicates if the breakthrough has been identified as a spillover. Source column additionally indicates if the spillover has been identified as part of an expert interview (I). Spillover sources are indicated in brackets and follow the taxonomy developed in \cite{Stephan_2021}: Other technology (T), other sector (S), other scientific discipline (D). }
    \label{tab:innovations}
\end{table}
